A SAT-Based Counterexample Guided Method for Unbounded Synthesis

### 2016-04-28 paper.tex
Example dump, needs lots of editing

This example is the $n=2$ instance of the more general problem of an arbiter of
$n$ resources. The set of winning states grows exponentially with the number of
resources the controller could use to grant requests, which makes the problem
hard for BDD solvers. This example illustrates the strength of the approach
presented here, which does not need to enumerate all winning states as will be
shown in Section~3.

@@

\tikzstyle{unfixed}=[dash pattern = on 2pt off 2pt]

@@

@@

@@

                node [above=4pt] {$\langle s, k \rangle$};
                        node [left=4pt] {\texttt{gr0=1, gr1=0}}
                    node [left=4pt] {\texttt{gr0=0, gr1=1}}

@@

                node [above=4pt] {$\langle s, k \rangle$};

@@

        \caption{AGT with partial strategy}
        \label{fig:strategy}

@@

                node [above=4pt] {$\langle s, k \rangle$};

@@

        \begin{tikzpicture}[scale=0.75,dash pattern = on 2pt off 2pt, level distance = 10mm]
                            edge from parent [fixed] node [left] {$c$}
                        edge from parent [fixed] node [left] {$b$}
                    edge from parent [fixed] node [left] {$a$}
                            edge from parent [fixed] node [left] {$f$}
                        edge from parent [fixed] node [right] {$e$}
                    edge from parent [fixed] node [right] {$d$}
        \caption{AGT extended to bound 3 with arbitrary controller actions}
        \label{fig:extended}

@@

Given an environment (controller) abstract game tree $T$ a \emph{partial
strategy} $Strat: \textsc{nodes}(T) \rightarrow C$ ($Strat: \textsc{nodes}(T)
\rightarrow U$) labels each node of the tree with the controller's
(environment's) action to be played in that node.   Given a partial strategy
$Strat$, we can map each leaf $l$ of the abstract game tree to $\langle
s',i'\rangle=\textsc{outcome}(\langle s, i\rangle, Strat, l)$ obtained by
playing all controllable and uncontrollable actions on the path from the root
to the leaf.  Figure~\ref{fig:strategy} shows an example partial strategy for
the controller.  The controller responds to the requests by first granting
access to \texttt{resource1}, then to \texttt{resource0} in the second round.

@@

\begin{exmp}
    The controller's partial strategy in
    Figure~\ref{fig:strategy} is defeated by the counterexample action
    \texttt{req=2}.  Figure~\ref{fig:refined} shows a refined AGT that includes
    this action.
\end{exmp}

>>>>

This example is the $n=2$ instance of the more general problem of an arbiter of
$n$ resources. The set of winning states grows exponentially with the number of
resources the controller could use to grant requests, which makes the problem
hard for BDD solvers. In Section~3 we will outline how this example can be
solved without enumerating all winning states.

@@

@@

        \begin{tikzpicture}[level distance = 5mm]
            \node [circle,draw,inner sep=1pt] (root){}
                child {node [circle,draw,inner sep=1pt] {}
                    child {node [circle,draw,inner sep=1pt] {}
                        child {node [circle,draw,inner sep=1pt] {}
                            child {node [circle,draw,inner sep=1pt] {}
                                child {node [circle,draw,inner sep=1pt] {}
                                    node [left=4pt] {\texttt{gr0=0 gr1=0}}
                                node [left=4pt] {\texttt{req=1}}
                            node [left=4pt] {\texttt{gr0=0 gr1=0}}
                        node [left=4pt] {\texttt{req=1}}
                node [left=4pt] {\texttt{gr0=0 gr1=0}}
                node [left=4pt] {\texttt{req=1}}
        \caption{Environment-winning trace}
        \label{fig:trace}
    \begin{subfigure}[t]{.1\textwidth}
        \begin{tikzpicture}[level distance = 5mm]
            \node [circle,draw,inner sep=1pt] (root){}
                child {node [circle,draw,inner sep=1pt] {}
                    child {node [circle,draw,inner sep=1pt] {}
                        child {node [circle,draw,inner sep=1pt] {}
                            child {node [circle,draw,inner sep=1pt] {}
                                child {node [circle,draw,inner sep=1pt] {}
                                    node [left=4pt] {\texttt{gr0=0 gr1=0}}
                                node [left=4pt] {\texttt{req=1}}
                            node [left=4pt] {\texttt{gr0=0 gr1=1}}
                        node [left=4pt] {\texttt{req=1}}
                node [left=4pt] {\texttt{gr0=0 gr1=0}}
                node [left=4pt] {\texttt{req=1}}
        \caption{Controller-winning trace}
        \label{fig:trace2}
    \begin{subfigure}[t]{.1\textwidth}

@@

                        % node [left=4pt] {\texttt{gr0=1 gr1=0}}

@@

                    % node [left=4pt] {\texttt{gr0=0 gr1=1}}

@@

@@

        \label{fig:agt2}

@@

                node [above=4pt] {$\langle s, 2 \rangle$};

@@

@@

Given an environment (controller) abstract game tree $T$ a \emph{partial
strategy} $Strat: \textsc{nodes}(T) \rightarrow C$ ($Strat: \textsc{nodes}(T)
\rightarrow U$) labels each node of the tree with the controller's
(environment's) action to be played in that node.   Given a partial strategy
$Strat$, we can map each leaf $l$ of the abstract game tree to $\langle
s',i'\rangle=\textsc{outcome}(\langle s, i\rangle, Strat, l)$ obtained by
playing all controllable and uncontrollable actions on the path from the root
to the leaf.

%%%Figure~\ref{fig:strategy} shows an example partial strategy for
%%%the controller.  The controller responds to the requests by first granting
%%%access to \texttt{resource1}, then to \texttt{resource0} in the second round.

\begin{exmp}

    To solve our example, the algorithm begins by invoking a SAT solver to find
    a trace of length $k$ from the initial state to the error state (see
    Figure~\ref{fig:trace}). This trace is used to create an abstract game tree
    by taking the first action in the trace and fixing it as the first action
    in the game tree (Figure~\ref{fig:agt}).  This abstract game tree
    represents the conjecture that \texttt{req=1} may be a good first action.
    Next the solver attempts to find a counterexample to this conjecture by
    finding a trace that avoids the error state with the first environment
    action fixed to \texttt{req=1} (Figure~\ref{fig:trace2}), i.e.  by playing
    the abstract game on behalf of the controller.

    Observe that the controller chooses to grant only one resource, just enough
    to ensure that an error state will not be reached. This example will show
    the worst case when the SAT solver returns actions that will lead to large
    abstract game trees. However, in practice we employ heuristics to guide the
    SAT solver away from the worst case.

    The environment's candidate is refined using the controller's
    counterexample traces. In Figure~\ref{fig:agt2}, the game tree tree is
    refined based on the first action of the controller's trace. The
    \texttt{req=1} action in the second round is still winning under the
    assumption that the controller will not grant any resources. The
    environment is free to make that assumption because the controller's
    counterexample does not yet contain any grants. Eventually, the refinements
    to the environment's candidate will force the controller to grant more
    resources and result in a game tree such as in Figure~\ref{fig:refined}.

@@

%%%\begin{exmp}
%%%    The controller's partial strategy in
%%%    Figure~\ref{fig:strategy} is defeated by the counterexample action
%%%    \texttt{req=2}.  Figure~\ref{fig:refined} shows a refined AGT that includes
%%%    this action.
%%%\end{exmp}
